---
title: "Gapminder ggplot 2"
author: "BM"
date: "September 27, 2016"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse) # gets me ggplot 2
library(gapminder)



ggplot(gapminder, aes(x = gdpPercap, y = lifeExp)) # will plot nothing - need to specify geom- the layer
p <- ggplot(gapminder, aes(x = gdpPercap, y = lifeExp))

## add points to plot by:

p + geom_point()
# in the graph- all data smushed on one side - need to log transform x
# brute force

ggplot(gapminder, aes(x = log10(gdpPercap), y = lifeExp)) + geom_point() # not right

# doing it right - scalling data to units of your figure

ggplot (gapminder, aes(x = gdpPercap, y = lifeExp)) + 
  scale_x_log10() +
  geom_point()

# a package is avail to change the numbers from scientific to another

## lets save this 
p <- ggplot (gapminder, aes(x = gdpPercap, y = lifeExp)) + 
  scale_x_log10() +
  geom_point()
p

P <- ggplot (gapminder, aes(x = gdpPercap, y = lifeExp)) + 
  scale_x_log10() 
P

  P + geom_point()
  
## color by continent
P + geom_point(aes(color = continent)) # mapping an aesthetic


## Deal with overplotting
P + geom_point(alpha = 1/3, size = 3) ## setting an aesthetic

#Above both take P and add points but in 50 line we are saying consult a data and map color to continent. for 54, we are saying for every point that is third, make the size this - has nothing to do with data

P + geom_point (aes(color = continent), alpha = 1/3, size = 3) # adding color

P + geom_point(alpha = 1/3, size = 3) # no color

#add a fitted line or curve

P+
  
geom_point(aes(color = continent))+
  geom_smooth(aes( color = continent)) # will create a legent for all countries

# better eg

P+
  
geom_point(aes(color = continent))+
  geom_smooth(aes( color = continent))


# you can  copy paste above - the color- this in your original p

P+
  geom_point(aes(color = continent))+
  geom_smooth(se = FALSE, methid ="lm") 


#facets - makes a mini plot for every level of continent


P + geom_point(alpha = 1/3, size = 3) +
  facet_grid( year ~ continent) # looks bad - grid lays them out in a grid.


P + 
  geom_point(alpha = 1/3, size = 3) +
  geom_smooth(se = FALSE)
    facet_wrap(~ continent)

    
    


